---
title: "CMA Exercise 2"
author: "Linus Rüegg"
date: "4/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times
```

```{r Data Import}
wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",")

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
wildschwein_BE
```

# Task 1
```{r Task 1: Timelag}
# Creating new column showing the timelag between two entries in int seconds

wildschwein_BE <- wildschwein_BE %>%
  group_by(TierID) %>%
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs")))

```

### How many individuals were tracked?
```{r Task 1: Names & occurences of individuals}
table(wildschwein_BE$TierName)
```
Three, all with above 14'000 tracker fixes.


### For how long were the individuals tracked? Are there gaps?
### Were all individuals tracked concurrently or sequentially?
### What is the temporal sampling interval between the locations?

```{r Task 1: Tracking Behavior}
wildschwein_BE %>%
  group_by(TierName) %>%
  summarise(min_time = min(DatetimeUTC),
            max_time= max(DatetimeUTC),
            Tracking_Period = difftime(max_time, min_time),
            min_lag = min(timelag, na.rm=T),
            max_lag = max(timelag, na.rm=T),
            median_lag = median(timelag, na.rm=T)
            )
```
The Tracking_Period for each of our three boars is a bit different.
Looking at min_time and max_time of the three, they were tracked concurrently.
And finally looking at the timelag: There is no strict sampling interval, sometimes the next location was sampled just 12s after the previous and the most extreme gap was 60'367s long. The median timelag for all three boars was around 15min (900s).

```{r Task 1: Plot}
ggplot(wildschwein_BE)+
  geom_point( aes(x = DatetimeUTC, y = timelag, color = TierName, shape = TierName))
```
The plot proves the assumptions done by just looking at the summary statistics.

# Task 2
```{r Task 2: Steplength}
wildschwein_BE <-  wildschwein_BE %>%
  group_by(TierName) %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
```

```{r Task 2: Speed!}
wildschwein_BE <- wildschwein_BE %>%
  group_by(TierName) %>%
  mutate(speed = steplength/timelag)
```
### What speed unit do you get?
m/s

#  Task 3
```{r Data Import 2}
caro <- read_delim("caro60.csv",",")

caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)
nrow(caro)
```

```{r Data Sequencing}
caro_3 <- caro[seq(1,nrow(caro),by=3),]
caro_6 <- caro[seq(1,nrow(caro),by=6),]
caro_9 <- caro[seq(1,nrow(caro),by=9),]
```

```{r}
nrow(caro_3)
nrow(caro_6)
nrow(caro_9)
```
```{r Adding timelag, steplength and speed}
# Caro_3
caro_3 <- caro_3 %>%
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs")))
caro_3 <- caro_3 %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
caro_3 <- caro_3 %>%
  mutate(speed = steplength/timelag)
# Caro_6
caro_6 <- caro_6 %>%
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs")))
caro_6 <- caro_6 %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
caro_6 <- caro_6 %>%
  mutate(speed = steplength/timelag)
# Caro_9
caro_9 <- caro_9 %>%
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs")))
caro_9 <- caro_9 %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
caro_9 <- caro_9 %>%
  mutate(speed = steplength/timelag)
# Caro
caro <- caro %>%
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs")))
caro <- caro %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
caro <- caro %>%
  mutate(speed = steplength/timelag)
```
### Plotting
```{r Calculating Line Geometries}
caro_l <- caro %>%
  group_by() %>%
    summarise(do_union = FALSE) %>%
      st_cast("LINESTRING")

caro_l_3 <- caro_3 %>%
  group_by() %>%
    summarise(do_union = FALSE) %>%
      st_cast("LINESTRING")

caro_l_6 <- caro_6 %>%
  group_by() %>%
    summarise(do_union = FALSE) %>%
      st_cast("LINESTRING")

caro_l_9 <- caro_9 %>%
  group_by() %>%
    summarise(do_union = FALSE) %>%
      st_cast("LINESTRING")
```


```{r Org/3Min}
ggplot() +
  ggtitle("Caro's Route - Original vs 3 Min resampled data")+
  xlab("E")+
  ylab("N")+
  theme_set(theme_minimal()) +
  theme(aspect.ratio=1) +
  geom_sf(data = caro, aes(color = "1 Min"), alpha = 0.8) +
  geom_sf(data = caro_l, aes(color = "1 Min"), alpha = 0.8) +
  geom_sf(data = caro_3, aes(color = "3 Min"), alpha = 0.8)+
  geom_sf(data = caro_l_3, aes(color = "3 Min"), alpha = 0.8)+
  coord_sf(datum = 2056) +
  scale_color_manual(name = "Sampling Interval", breaks = c("1 Min", "3 Min"), values = c("1 Min" = "cyan", "3 Min" = "darkgreen"))
```

```{r Org/6Min}
ggplot() +
  ggtitle("Caro's Route - Original vs 6 Min resampled data")+
  ylab("N")+
  xlab("E")+
  theme_set(theme_minimal()) +
  theme(aspect.ratio=1) +
  geom_sf(data = caro, aes(color = "1 Min"), alpha = 0.8) +
  geom_sf(data = caro_l, aes(color = "1 Min"), alpha = 0.8) +
  geom_sf(data = caro_6, aes(color = "6 Min"), alpha = 0.8)+
  geom_sf(data = caro_l_6, aes(color = "6 Min"), alpha = 0.8)+
  coord_sf(datum = 2056) +
  scale_color_manual(name = "Sampling Interval", breaks = c("1 Min", "6 Min"), values = c("1 Min" = "cyan", "6 Min" = "blue"))
```
```{r Org/9Min}
ggplot() +
  ggtitle("Caro's Route - Original vs 9 Min resampled data")+
  ylab("N")+
  xlab("E")+
  theme_set(theme_minimal()) +
  theme(aspect.ratio=1) +
  geom_sf(data = caro, aes(color = "1 Min"), alpha = 0.8) +
  geom_sf(data = caro_l, aes(color = "1 Min"), alpha = 0.8) +
  geom_sf(data = caro_9, aes(color = "9 Min"), alpha = 0.8)+
  geom_sf(data = caro_l_9, aes(color = "9 Min"), alpha = 0.8)+
  coord_sf(datum = 2056) +
  scale_color_manual(name = "Sampling Interval", breaks = c("1 Min", "9 Min"), values = c("1 Min" = "cyan", "9 Min" = "darkblue"))
```
I do not like the uneven grid spacing. Will maybe fix this when trying to flip the plot again.

```{r Speed with different Sampling Interval}
ggplot()+
  ggtitle("Speed with different Sampling Interval")+
  ylab("Speed [m/s]")+
  xlab("Time [s]")+
  theme_set(theme_minimal()) +
  geom_line(data = caro, aes(x = DatetimeUTC, y = speed, color ="1 Min"))+
  geom_line(data = caro_3, aes(x = DatetimeUTC, y = speed, color ="3 Min"))+
  geom_line(data = caro_6, aes(x = DatetimeUTC, y = speed, color ="6 Min"))+
  geom_line(data = caro_9, aes(x = DatetimeUTC, y = speed, color ="9 Min"))+
  scale_color_manual(name = "Sampling Interval", breaks = c("1 Min","3 Min", "6 Min","9 Min"), values = c("1 Min" = "cyan", "9 Min" = "darkblue", "6 Min" = "blue", "3 Min" = "darkgreen"))
```
# Task 4

```{r Zoo familiarisation}
library(zoo)

example <- rnorm(10)
rollmean(example,k = 3,fill = NA,align = "left")
##  [1]  0.93634335  0.31709038  0.02370048  0.67869801  0.73369105  0.50401344
##  [7] -0.56144365 -0.56902598          NA          NA
rollmean(example,k = 4,fill = NA,align = "left")
##  [1]  0.6775521  0.2045005  0.5848215  0.5255629  0.3446928  0.1459635
##  [7] -0.4102301         NA         NA         NA
```
```{r Rolling Window on Caro data}
carorw <- caro %>%
  mutate(k3 = rollmean(speed, k= 3, fill = NA, align = "left"),
          k6 = rollmean(speed, k= 6, fill = NA, align = "left"),
          k15 = rollmean(speed, k= 15, fill = NA, align = "left"),
          k30 = rollmean(speed, k= 30, fill = NA, align = "left"))
```

```{r PLotting the Rolling Window data}
ggplot(data = carorw)+
  ggtitle("Rolling Window Caro Data")+
  ylab("Speed [m/s]")+
  xlab("Time [t]")+
  theme_set(theme_minimal()) +
  geom_line(aes(x = DatetimeUTC, y = speed, color ="Original"))+
  geom_line(aes(x = DatetimeUTC, y = k3, color ="k 3"))+
  geom_line(aes(x = DatetimeUTC, y = k6, color ="k 6"))+
  geom_line(aes(x = DatetimeUTC, y = k15, color ="k 15"))+
  geom_line(aes(x = DatetimeUTC, y = k30, color ="k 30"))+
  scale_color_manual(name = "Window Size", breaks = c("Original","k 3", "k 6","k 15", "k 30"), values = c("Original" = "cyan", "k 15" = "purple", "k 6" = "darkgreen", "k 3" = "blue", "k 30" = "darkred"))
```

